---
title: "Statistical  Inference Assignment"
author: "Carey Raychelle"
date: "2023-06-19"
output: pdf_document
---

##Assignment Description
Investigating exponential distribution in R and comparing with the Central Limit Theorem.
Exponential distribution in R can be simulated by rexp(n,lamda) where lamda is rate.Mean and standard deviation of distribution is 1/lamda Lamda is set to be 0.2
```{r}
#install the packages needed
library(knitr)
library(ggplot2)
#making the report reproducible
set.seed(12345)
```
#Simulation Exercise
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

```{r}
#setting variations for the simulation
n <- 40
lamda <- 0.2
sim <- 1000
#Create a matrix with 1000 rows and columns 40 of random simulations
matsim <- matrix(rexp(sim * n,rate = lamda),sim,n)
#Calculate the means and plot
sim_mean <- rowMeans(matsim)
hist(sim_mean, xlab="Mean of 40 exponentials",ylab="Frequency",main="Histogram of Simulation Mean",col="purple")
```
#Sample Mean and Theoretical Mean
```{r}
sample_mean <- mean(sim_mean)
sample_mean
```
```{r}
theoritical_mean <- 1/lamda
theoritical_mean
```
From the above the sample mean which is 4.97192 is very close to the theoretical mean which is 5

#Sample Variance and Theoretical Variance
```{r}
sample_var <- var(sim_mean)
sample_var
```
```{r}
theoretical_var <- (1/lamda)^2/n
theoretical_var
```
The sample variance is also close to the theoretical variance

##Approximate Normal Distribution
Creating an approximate normal to see how the sample differs from it
```{r}
plotnorm <- data.frame(sim_mean)
g <- ggplot(plotnorm,aes(x=sim_mean))
g=g+geom_histogram(aes(y=after_stat(density)),colour="green",fill="orange")
g=g+geom_density(colour="black", size=1)
g=g+ggtitle("Histogram of Simulation Mean")
g
```
The plot indicates that the histogram can be approximated with the normal distribution
#Comparing Confidence Intervals
```{r}
sample_conf <- round(mean(sim_mean)+c(-1,1)*1.96*sd(sim_mean)/sqrt(n),3)
sample_conf
```
```{r}
theoretical_conf <- round(mean(theoritical_mean)+c(-1,1)*1.96*sqrt(theoretical_var)/sqrt(n),3)
theoretical_conf
```
The sample 95% confidence interval [4,729,5.215] is close to the theoretical 95% confidence interval of [4.755,5.245]

Q-Q Plot for Quantiles
```{r}
qqnorm(sim_mean,main = "Normal Q-Q Plot", xlab="Theoretical Quantiles",ylab="Sample Quantiles")
qqline(sim_mean,col="blue")
```
Distribution is approximately normal
